---
title: "exercise 2 - Bivariate Case"
output: html_document

# Zhiyi Wang
---

```{r, include=FALSE}
link="https://github.com/EvansDataScience/data/raw/master/crime.RData"
load(file = url(link))

names(crime)

head(crime)

str(crime,width = 70,strict.width='cut')
```

```{r table_rel_PIPES,eval=FALSE}
library(magrittr)
(CrimeTotal=table(crime$crimecat,crime$Occurred.DayTime)%>% 
        prop.table() %>% 
        "*"(100)%>% 
        round(2) 
        )

CrimeCol=table(crime$crimecat,crime$Occurred.DayTime)%>%
         prop.table(margin = 2)%>%  
         "*"(100)%>%
         round(3)

df.T=as.data.frame(CrimeTotal)
head(df.T)

names(df.T)=c('Crime','Daytime','Percent') 
head(df.T)

library(ggplot2)                           
base = ggplot(df.T, aes(Daytime,Crime)) 

tablePlot1 = base + geom_point(aes(size = Percent), colour = "gray") 

tablePlot2 = tablePlot1 + geom_text(aes(label = Percent),
                                    nudge_x = 0.1,
                                    size=2)

tablePlot3 = tablePlot2 + scale_size_continuous(range=c(0,10)) 
tablePlot4 = tablePlot3 + theme_minimal() 
tablePlot4 + theme(legend.position="none") 
```

```{r facet, eval=FALSE}
base  = ggplot(df.T, aes(x = Crime, y = Percent ) ) 
bars1 = base + geom_bar( stat = "identity" ) + theme_minimal()
bars2 = bars1 + facet_wrap( ~ Daytime ,nrow = 1) 
bars3 = bars2 + theme( axis.text.x = element_text(angle = 90,
                                                  hjust = 1,
                                                  size=3 ) )
bars3
```

```{r flip_facet, eval=FALSE}
df.C=as.data.frame(CrimeCol)
colnames(df.C)=c('Crime','Daytime','Percent')

base  = ggplot(df.C, aes(x = Crime, y = Percent ) ) 
bars1 = base + geom_bar( stat = "identity" )
bars2 = bars1 + facet_wrap( ~ Daytime ,nrow = 1) 
bars2 + coord_flip()
```

```{r orderFacet, eval=FALSE}
base  = ggplot(df.C, aes(x = reorder(Crime, Percent), y = Percent ) ) 
bars1 = base + geom_bar( stat = "identity" )
bars2 = bars1 + facet_wrap( ~ Daytime ,nrow = 1) 
bars2 + coord_flip() + theme(axis.text.y = element_text(size=4,angle = 45)) 
```

#exercise-1:Turn the bars into lollipop with the right components.

```{r, eval=FALSE}
titleText1='Exercise 1: Crime rate in different time zones'
sourceText='Source: Seattle Open Data'
base  = ggplot(df.C, aes(x = reorder(Crime, Percent), y = Percent ) ) 
lolliplot1 = base + geom_bar( stat = "identity" )
lolliplot2 = lolliplot1 + facet_wrap( ~ Daytime ,nrow = 1) 
lolliplot2 + coord_flip() + theme(axis.text.y = element_text(size=4,angle = 45)) + labs(title=titleText1,
                     caption = sourceText) 
```


#Exercise-2:Change the heatplot to ascending order, where intensity goes from yellow to purple.

```{r heatDescending, eval=FALSE}
# heatplot
titleText2='Exercise 2: Crime rate in different time zones (heatplot)'
base  = ggplot(df.C, aes(x = Daytime, y = reorder(Crime, Percent), fill = Percent)) 
heat1 = base +  geom_tile() 
heat2 = heat1 +scale_fill_gradient(low = "yellow", 
                                   high = "purple")
heat3 = heat2 + theme_classic() 

heat4 = heat3 + theme(axis.text.x = element_text(angle = 90, vjust = 0.6), 
                      legend.title = element_blank(), 
                      legend.position="top", 
                      legend.direction="horizontal",
                      legend.key.width=unit(1, "cm"),
                      legend.key.height=unit(1, "cm")) 

heat4 + labs(y="Crime", title=titleText2, caption = sourceText)
```

## Categorical-Numerical relationships

```{r, eval=FALSE}
summary(crime$DaysToReport)
tapply(crime$DaysToReport,crime$Precinct,mean)
boxplot(DaysToReport~Precinct,data=crime)
tapply(crime$DaysToReport,crime$Precinct,summary)
```

```{r, eval=FALSE}
boxplot(DaysToReport~Precinct,data=crime,
        subset = DaysToReport>1) 

par(mfrow=c(2,2)) 
boxplot(DaysToReport~Precinct,data=crime,subset = DaysToReport>=7,
        main="One week or more")

boxplot(DaysToReport~Precinct,data=crime,subset = DaysToReport>=30,
        main="One 30-day month or more")

boxplot(DaysToReport~Precinct,data=crime,subset = DaysToReport>=180,
        main="180 days or more")

boxplot(DaysToReport~Precinct,data=crime,subset = DaysToReport>=365,
        main="One year or more")
```

```{r, eval=FALSE}
crimeYear=crime[crime$DaysToReport>=365,]
tapply(crimeYear$DaysToReport,crimeYear$Precinct,length)
tapply(crimeYear$DaysToReport,crimeYear$year,length)
crimeY2000=crime[(crime$DaysToReport>=365) & (crime$year>=2000),]
tapply(crimeY2000$DaysToReport,crimeY2000$Precinct,length)
```

```{r boxpAfter2000, eval=FALSE}
boxplot(DaysToReport~Precinct,data=crimeY2000,
        main="One year or more (from 2000)")

crimeY2000$YearsToReport=crimeY2000$DaysToReport/365
boxplot(YearsToReport~Precinct,data=crimeY2000,
        main="One year or more (from 2000)")

boxplot(YearsToReport~Precinct,data=crimeY2000,subset = YearsToReport>=5,
        main="Five years or more (from 2000)")
```

```{r exploreCatBOX, eval=FALSE}

par(mfrow=c(2,1))

boxplot(YearsToReport~weekday,data=crimeY2000,
        main="One year or more BY WEEKDAY (from 2000)",las=2)

boxplot(YearsToReport~Occurred.DayTime,data=crimeY2000,
        main="One year or more BY TIME CRIME OCCURRED (from 2000)",las=2)

```

```{r exploreBOX2, eval=FALSE}
par(mfrow=c(2,1))

boxplot(YearsToReport~year,data=crimeY2000,
        main="One year or more (from 2000)",las=2)

boxplot(YearsToReport~crimecat,data=crimeY2000,
        main="One year or more (from 2000)",las=2)
```

```{r yearsToRep, eval=FALSE}
crimeYearGG=crimeY2000[complete.cases(crimeY2000$YearsToReport),]

base = ggplot(crimeYearGG,aes(x=factor(year), y=YearsToReport)) 
box1  = base + geom_boxplot()
box2  = box1 + labs(title="exercise3:One year or more (from 2000)",
                     caption = "the Seattle Open Data portal")
box3 = box2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
box3

```

```{r, eval=FALSE}
base  = ggplot(crimeYearGG,aes(x=factor(year), y=YearsToReport))
box1  = base + geom_boxplot()
box2  = box1 + labs(title="exercise3:One year or more (from 2000)",
                     caption = sourceText)
box3 = box2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
mins = box3 + stat_summary(fun.y=min, 
                           geom="line",
                           show.legend = T,size=1,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",
                              linetype='dashed',
                              size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median,
                                    geom="line",size=2,
                                    aes(group=1,col='Median'))
all1=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red")
                                      )
all1
```

```{r, eval=FALSE}
crimeWeek=crime[(crime$DaysToReport<365) & (crime$year>=2000),]
crimeWeek$WeeksToReport=crimeWeek$DaysToReport/7
crimeYearGG2=crimeWeek[complete.cases(crimeWeek$WeeksToReport) &complete.cases(crimeWeek$crimecat),]

base = ggplot(crimeYearGG2,aes(x=factor(year), y=WeeksToReport)) 
base1  = base + geom_boxplot()
base2  = base1 + labs(title="exercise3:The crimes that took less than a year to report since 2000",
                     caption = sourceText)
base3 = base2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))

mins = base3 + stat_summary(fun.y=min,size=1,
                           geom="line", linetype='dashed',show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median,
                                    geom="line",size=2,
                                    aes(group=1,col='Median'))
all2=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red")
                                      )
all2 
```

```{r, eval=FALSE}
base= ggplot(crimeYearGG,
             aes(x = reorder(crimecat, YearsToReport, FUN = max), 
                 y=YearsToReport)) 
base1  = base + geom_boxplot()
base2  = base1 + labs(title="exercise3:The crimes that took less than a year to report since 2000",
                     caption = sourceText)
base3 = base2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
mins = base3 + stat_summary(fun.y=min,size=1,
                           geom="line", linetype='dashed',show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median, size=2,
                                    geom="line",
                                    aes(group=1,col='Median'))
all3=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red"))
all3 + coord_flip()
```

```{r byCrime2, eval=FALSE}
base = ggplot(crimeYearGG2,
              aes(x = reorder(crimecat, WeeksToReport, FUN = max),
                  y=WeeksToReport)) 
base1  = base + geom_boxplot()
base2  = base1 + labs(title="exercise3:The crimes that took less than a year to report since 2000",
                     caption = sourceText)
base3 = base2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
mins = base3 + stat_summary(fun.y=min,size=1,
                           geom="line", linetype='dashed',show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median,size=2,
                                    geom="line",
                                    aes(group=2,col='Median'))
all3=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red"))
all3+coord_flip()
```

#Exercise-3:Complete the information needed in the previous plots.

```{r, eval=FALSE}
anovaData=crimeY2000[crimeY2000$YearsToReport>=5,]

tapply(anovaData$YearsToReport, anovaData$Precinct, mean,na.rm=T)

install.packages("Rmisc")
library(Rmisc)

group.CI(YearsToReport ~ Precinct, 
         data=anovaData, 
         ci = 0.95)
```

```{r, eval=FALSE}
anovaData=anovaData[complete.cases(anovaData),]

install.packages("ggpubr")
library(ggpubr)
ggline(data=anovaData,x = "Precinct", y = "YearsToReport",add = 'mean_ci',
       error.plot = "pointrange") + scale_y_continuous(breaks=seq(7,10,0.5))

```

```{r testAnova, eval=FALSE}
res.aov <- aov(YearsToReport ~ Precinct, data = anovaData)

summary(res.aov)[[1]]$Pr[1]

kruskal.test(YearsToReport ~ Precinct, data = anovaData)
```

## Numerical-Numerical relationships

```{r, eval=FALSE}
daysByNeigh=tapply(crime$DaysToReport, crime$Neighborhood, mean,na.rm=T)
head(daysByNeigh)

crimesByNeigh=tapply(crime$crimecat, crime$Neighborhood, length)%>%      
                     prop.table()%>%
                     "*"(100)%>% 
                     round(2) 
head(crimesByNeigh)

library(tibble)
as.data.frame(daysByNeigh)%>%rownames_to_column()

daysByNeigh=as.data.frame(daysByNeigh)%>%rownames_to_column()
crimesByNeigh=as.data.frame(crimesByNeigh)%>%rownames_to_column()

num_num=merge(daysByNeigh,crimesByNeigh) 
head(num_num)

base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh)) 
plot1= base +  geom_point() 
plot2 = plot1 + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood",
                     caption =sourceText)
plot3 = plot2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
plot3
```

```{r ggscatter, eval=FALSE}
library(ggrepel)
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh,
                           label=rowname)) 
plot1= base +  geom_point() 
plot2 = plot1 + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood",
                     caption =sourceText)
plot3 = plot2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
plot3 + geom_text_repel()
```

```{r, eval=FALSE}
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh,label=rowname)) 
plot1= base +  geom_point() 
plot2 = plot1 + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood",
                     caption =sourceText)
plot3 = plot2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
plot3 + geom_text_repel(aes(label=ifelse(crimesByNeigh>=5,
                                         num_num$rowname, "")))
```

```{r simpleScatter,eval=FALSE}
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh)) 
scatp1 = base +  geom_point() 
scatp2 = scatp1 + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood",
                     caption =sourceText)
scatp3 = scatp2 + theme(panel.background =element_rect(fill = "white",colour = "grey50"))
scatp3 + geom_text(aes(label=ifelse(crimesByNeigh>=5,num_num$rowname, "")))
```

```{r scatterSEARCH, eval=FALSE}
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh,label=rowname)) 
base +  geom_point() + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood",
                     caption =sourceText) + theme(panel.background =element_rect(fill = "white",colour = "grey50")) +
geom_text_repel(aes(label=ifelse(num_num$rowname=='NORTHGATE',
                                                        num_num$rowname, "")))
```

```{r hexbins, eval=FALSE}
install.packages("hexbin")
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh)) 
scatp1 = base +  geom_hex(bins = 10)
scatp2= scatp1 + geom_text_repel(aes(label=ifelse(crimesByNeigh>=5,
                                                  num_num$rowname,
                                                  ""))) 
scatp2 + scale_fill_distiller(palette ="Greys",direction=1) + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood (at least 5%)",
                     caption =sourceText)

```

```{r density,eval=FALSE}
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh)) 
scatp1 = base +  stat_density_2d(aes(fill = ..density..), 
                                 geom = "raster", contour = FALSE)
scatp2=scatp1+geom_text_repel(aes(label=ifelse(crimesByNeigh>=5,
                                               num_num$rowname, "")))
scatp3 = scatp2 +  theme(legend.position='none') 
scatp4= scatp3 + scale_fill_distiller(palette="Greys", direction=1)  + labs(title="exercise4:Means of days to report and proportion of crimes by neighborhood (at least 5%)",
                     caption =sourceText)

scatp5 = scatp4 +  scale_x_continuous(expand = c(0, 0)) + 
         scale_y_continuous(expand = c(0, 0)) 
scatp5
```

#Exercise-4:Complete the elements missing in the previous plots.

